Developing a preprint culture in biology

### 2012-10-05 ms.tex
merge

Public preprint servers allow authors to make manuscripts publicly available
before, or in parallel to, submitting them to journals for traditional
peer-review. The rationale for preprint servers is fundamentally simple: to make
the results available to the scientific community as soon as possible rather
than to wait until the peer-review process is fully completed. The goal of arXiv
and open preprint servers is not to circumvent the peer-review process; almost all
manuscripts on arXiv are submitted to peer-review.  Sharing ideas as quickly as
possible using preprint servers has numerous advantages.  These include rapid
dissemination of work-in-progress to a wider audience, including readers in
developing countries where access to subscription journals is often a limiting
factor in scientific research.  Furthermore, preprints provide the opportunity
to solicit feedback from a larger pool of reviewers and can be seen as an
integral aspect of a vigorous peer-review process \cite{hoc12}.  Preprint
servers therefore increase the number of opportunities for review and revision
prior to publication, resulting in higher quality submissions and also could
alleviate reviewer burden.

Preprints gained popularity 20 years ago with the advent of arXiv, an open
preprint server widely used in physics and mathematics \cite{nie08,gin11}. Preprints
are also integral to the culture of other scientific fields.  Paul Krugman noted
that, in economics, the \emph{traditional model of submit, get refereed,
publish, and then people will read your work broke down a long time ago. In
fact, it had more or less fallen apart by the early 80s} \cite{kru12}. In
addition to a section in arXiv, economists have also the RePEc (Research Papers
in Economics) initiative, which aims to create an archive of working papers,
manuscripts, and book chapters.  Despite the success of this approach in other
fields, most manuscripts in biology are not submitted to a public archive before
being submitted to peer-review. In this article, we first highlight the
advantages of open preprint servers for both scientists and publishers. We
then debunk a few misconceptions, discuss the policies of major publishers in
biology, and briefly review the most popular open preprint servers currently available.

@@

Furthermore, the review process as a whole is critically over-loaded, because
the number of active scientists increases, because the pressure to publish
increases, and because of an effect dubbed ``the tragedy of the reviewers
commons'' \cite{hoc09}.  At the same time, rejection rates are high in most journals
\cite{aar08,roh09}, and when not invited to submit a revision, authors must start the whole
process all over again.  Initiatives to reduce time from submission to
publication have emerged across the scientific community. Rohr et al.
\cite{roh09} called for the recycling and reuse of peer-reviews:
by attaching previous reviews and detailed replies to a new submission, both
the editor and the referees can gauge the work done on the manuscript, and
perhaps evaluate it with less prejudice. In a similar way, the \emph{Peerage
of Science} initiative allows authors to seek anonymous pre-review by their
peers. Some journals now accept to publish papers which received good
evaluations, with \emph{Animal Biology} having recently accepted first a paper
reviewed entirely with the \emph{Peerage of Science} \cite{abb12}, effectively
outsourcing the review process. A widespread use of preprint servers can
achieve the same goal of reducing the time spent in review. By putting a
manuscript for open comments and criticisms, the authors will receive valuable
feedback and can improve the version which will be submitted. With a rich
enough community of scientists depositing preprints, and commenting on them,
the process of an open pre- review can become widespread and will overall
increase the quality of first submissions.

\begin{figure}[ht!] \centering\includegraphics[width=0.50\textwidth]{map.pdf}
\caption { It can take several months, and even a few years, before a submitted
paper is officially published and citable.  The average time to publication
varies greatly between journals and can be as low as 104 days (Evolution for
2011) to 213 (PLOS One in 2010).  Meanwhile, few people are aware of the
research that has been done since, typically, only close colleagues are given
access to the preprints. With public preprint servers, the science is
immediately available and can be openly discussed, analyzed, and integrated into
current research. It benefits both science and publishers. Both want the papers
to be well-known and cited, and public preprints make it possible to integrate
research even before publication, greatly improving immediacy.  }
\label{fig:map} \end{figure}
% Dom : Do you have a reference for the stats in the legend?

Public preprint servers offer a much fairer way to establish intellectual
priority by making the work available when done. Some manuscripts will spend
much more time in the review process than others. Surprisingly, there is a
perception in biology that public preprints make it easier to steal ideas, as if
a scientific idea was only complete once published in a peer-reviewed journal
\cite{gin11}. Ironically, mathematicians and physicists have embraced arXiv in
part to establish priority in a fair way \cite{gin11,cal12}.
% Dom: That's a very interesting statement, if not the best one to justify preprints. Could you develop just a bit more the argument?

Prepublication reviews by a small network of colleagues is an important part of
the scientific process, attested by the fact that nearly all published papers
acknowledge comments by people not listed as co-authors.  Preprint servers
simply offer a way to extend this network of colleagues to the entire scientific
community. It ensures that science is not constrained by small networks of
scientists exchanging ideas.  Paul Ginsparg created arXiv.org in part for
democratic reasons: he wanted everyone from students in small universities to
Ivy-League professors to have access to the most recent scientific \emph{ideas}.
Ginsparg's revolutionary idea was simply to use the power of the internet for
preprints, not just for the end product, so the scientific process can be open
as soon as possible.
% Dom: I do accept this argument, although I have doubts it makes any difference. Just think about the number of comments posted on PLoS One...

@@

Submitting to a preprint server is becoming more common in biology, even though
it still involves a minority of papers. The quantitative biology section in
arXiv is experiencing faster growth in submissions than any other field
\cite{cal12}. Most scientific journals are preprint-friendly: Nature, PLOS, BMC,
PNAS, Science (mostly) \ref{table:policies}, and all the journals from Elsevier %something is off in the formatting here
and Springer.  The Ecological Society of America and the Genetics Society of
America recently changed their policies to allow public preprints.  Few
scientific publications will not consider a manuscript submitted to a public
preprint server.  Still, a few journals adopt a ``by default'' hostile attitude
towards preprints, mostly due to the lack of clear policy of the publishers, or
perhaps because a preprint culture has not developed in biology and the practice
is still considered unusual. As an example, Wiley-Blackwell, which publishes
some leading journals, has no official policy on the subject
\ref{table:policies}.

Part of the hostility to preprints servers comes from a certain interpretation
of the Ingelfinger rule: scientists should not publish the same manuscript twice
\cite{alt96}. However, papers submitted to arXiv or other preprints servers are
only published in one peer-reviewed journal. A preprint is just a preprint, it
is a document that allows ideas to spread and be discussed, but it is not yet
formally validated by the peer-review system, which is why the majority of
publishers does not see arXiv and similar services as a violation of the
Ingelfinger rule. \emph{Nature} responded to the rumour that they refused
manuscripts submitted to arXiv by saying that ``\emph{Nature} never wishes to
stand in the way of communication between researchers. We seek rather to add
value for authors and the community at large in our peer review, selection and
editing'' \cite{nat05}.

@@

One of the biggest advantage of figshare over arXiv is that is it not limited to
quantitative sciences. arXiv.org has sections on quantitative biology but might
not be appropriate for non-quantitative work. With its flexible approach to
preprints, figshare offers an important alternative to arXiv for empirical
biologists. Furthermore, by allowing all types of content, figshare arguably
provides an archive for early results (e.g.: figures, lab presentations).

@@

F1000Research is not a public preprint server like the previous three servers.
Whereas arXiv, figshare, and PeerJ offer an option to submit a manuscript
without having it reviewed, papers submitted to F1000Research will eventually be
reviewed. Thus, F1000Research offers a hybrid model with publicly available
manuscripts at time of submission and standard peer-reviews. Manuscripts are
considered ``accepted'' and will only be indexed after two positive referee
response. F1000Research works closely with data providers to integrate raw data
to the paper. For instance, upon submitting a paper, authors are asked to upload
their data, which are then integrated in \emph{e.g.} figshare widgets, the DOI
of which are given in the paper when the data are first mentionned.  The
licensing of the data is similar to the one used by figshare, meaning that the
articles are free to access, and can be redistributed readily. By putting much
effort in integrating data to the paper, F1000Research is working to make
science more reproducible and open.

>>>>

Public preprint servers allow authors to make manuscripts publicly available
before, or in parallel to, submitting them to journals for traditional
peer-review. The rationale for preprint servers is fundamentally simple: to make
the results of scientific research available to the scientific community as soon
as possible, instead of waiting until the peer-review process is fully completed.
Sharing manuscripts using preprint servers has numerous advantages including: 1) rapid
dissemination of work-in-progress to a wider audience; 2) improved peer review by
encouraging feedback from the entire research community;
and 3) a fair and straightforward way to establish precedence.

Preprints began to gain popularity 20 years ago with the advent of arXiv, an open
preprint server widely used in physics and mathematics \cite{gin11}. Preprints
are also integral to the culture of other scientific fields.  Paul Krugman noted
that, in economics, the \emph{traditional model of submit, get refereed,
publish, and then people will read your work broke down a long time ago. In
fact, it had more or less fallen apart by the early 80s} \cite{kru12}. In
addition to a section in arXiv, economists have also the RePEc (Research Papers
in Economics) initiative, which aims to create an archive of working papers,
manuscripts, and book chapters.

Despite the success of this approach in other fields, most manuscripts in
biology are not posted to preprint servers and are therefore not seen by more
than a handful of other scientists prior to publication. In this article, we: 1)
highlight the advantages of open preprint servers for both scientists and
publishers; 2) address several misconceptions about preprints that are common
among biologists; 3) discuss the preprint policies of major publishers in
biology; and 4) review the most popular preprint servers currently available.

@@

Posting manuscripts as preprints also has the potential to result in higher
quality science, by allowing for pre-publication feedback
from a large pool of reviewers. Prepublication reviews by a small network of
colleagues are common in the biological sciences and are an important part of
the scientific process. These "friendly" reviews increases the chances of errors
being caught prior to publication, allow errors to be identified
prior to spending months in review, and allows authors to improve how they
communicate the results of their research by modifying their writing in response
to early feedback. Preprint servers offer a way to extend this network of
colleagues to the entire scientific
community, increasing the number of opportunities for review and revision
prior to publication, and resulting in higher quality submissions and publications.
As such, preprints can be seen as an
integral aspect of a vigorous peer-review process \cite{hoc12}.

%It is unclear to me whether we should keep this full paragraph or try to distill
%it to a sentence or two at the end of the previous paragraph.
Furthermore, the formal peer-review process as a whole is critically over-loaded. As
the number of active scientists increases and the pressure to publish
increases, it has become increasingly difficult for journals to find
reviewers \cite{hoc09}.  At the same time, rejection rates are high in most journals
\cite{aar08,roh09}, and when not invited to submit a revision, authors must start the
process over again at another journal. As a result, initiatives to reduce time from submission to
publication have emerged across the scientific community. Rohr et al.
\cite{roh09} called for the recycling and reuse of peer-reviews:
by attaching previous reviews and detailed replies to a new submission, both
the editor and the referees can gauge the work done on the manuscript, and
perhaps evaluate it with less prejudice. In a similar way, the \emph{Peerage
of Science} initiative allows authors to seek anonymous pre-review by their
peers. Some journals now accept to publish papers which received good
evaluations, with \emph{Animal Biology} having recently accepted first a paper
reviewed entirely with the \emph{Peerage of Science} \cite{abb12}, effectively
outsourcing the review process. A widespread use of preprint servers can
achieve the same goal of reducing the time spent in review. By allowing
open comments and criticisms prior to submission, the authors will receive valuable
feedback and can improve the version which will be submitted. With a rich
enough community of scientists depositing preprints, and commenting on them,
the process of an open pre-review can become widespread and will overall
increase the quality of first submissions.

\begin{figure}[ht!] \centering\includegraphics[width=0.50\textwidth]{map.pdf}
\caption { It can take several months, and even a few years, before a submitted
paper is officially published and citable.  The average time to publication
varies greatly between journals and can be as low as 104 days (Evolution for
2011) to 213 (PLOS One in 2010).  Meanwhile, few people are aware of the
research that has been done since, typically, only close colleagues are given
access to the preprints. With public preprint servers, the science is
immediately available and can be openly discussed, analyzed, and integrated into
current research. It benefits both science and publishers. Both want the papers
to be well-known and cited, and public preprints make it possible to integrate
research even before publication, greatly improving immediacy.  }
\label{fig:map} \end{figure}

Finally, public preprint servers offer a fair way to establish intellectual
priority by making the work available when immediately after it is complete.
Some manuscripts will spend much more time in the review process and/or in
production after acceptance, than others. This means that
that publication and acceptance dates do not accurately characterize who
came up with an idea first. For this reason, mathematicians and physicists
have embraced arXiv in part to establish priority in a fair way \cite{gin11,cal12}.

@@

In contrast to other disciplines, the field of biology has effectively no preprint
culture, with the exception of small pockets of primarily highly quantitative
research (e.g., epidemiology). While, submitting to preprint servers has become
more common in the past few years, and the quantitative biology section in arXiv
is experiencing rapid growth in submissions \cite{cal12}, the number of biology
papers submitted to preprint servers still represents only a small fraction of
the total research produced in biology.

There are a number of reasons that biologists have not traditionally had a
culture of sharing preprints, many of which are based on common misconceptions
regarding the costs and benefits of sharing preprints. For example, in contrast
to other fields there is a perception in biology that public preprints
make it easier to steal ideas. However, there is no evidence of this happening
in the numerous other fields that have adopted preprint servers, and since preprint
servers create a clear record of who had the idea first, and when, this
appears to be a largely unfounded concern. In other fields preprints serve the
opposite role, they allow straightforward establishment of precendence, letting
researchs lay claim to an idea thus prevening it from being "stolen".

Another major concern is that if a researcher posts a public preprint that they
will not be allowed to submit their paper to their journal of choice. This is 
based on a certain interpretation
of the Ingelfinger rule: scientists should not publish the same manuscript twice
\cite{alt96}. However, preprints are not peer-reviewed papers and posting to a
public server does not constitute publication in a peer-reviewed journal.
A preprint is simply a document that allows ideas to spread and be discussed, 
but it is not yet formally validated by the peer-review system. This is why the
majority of publishers do not see arXiv and similar services as a violation of the
Ingelfinger rule. Almost all of the major publishers in biology are preprint-friendly,
including: Nature Publishing Group, PLOS, BMC, PNAS, Science (mostly) \ref{table:policies},
and all the journals from Elsevier %something is off in the formatting here
and Springer.  The Ecological Society of America and the Genetics Society of
America also recently changed their policies to allow public preprints.  Few
scientific publications will not consider a manuscript submitted to a public
preprint server. In fact, \emph{Nature} responded to the rumour that they refused
manuscripts submitted to arXiv by saying that ``\emph{Nature} never wishes to
stand in the way of communication between researchers. We seek rather to add
value for authors and the community at large in our peer review, selection and
editing'' \cite{nat05}. Still, a few journals adopt a ``by default'' hostile attitude
towards preprints, mostly due to the lack of clear policy of the publishers, or
perhaps because a preprint culture has not developed in biology and the practice
is still considered unusual. As an example, Wiley-Blackwell, which publishes
some of the leading journals in biology, has no official policy on the matter.

\ref{table:policies}.

@@

One of the biggest advantage of figshare over arXiv is that is it not limited to
quantitative sciences. arXiv.org has sections on quantitative biology but might
not be appropriate for non-quantitative work. With its flexible approach to
preprints, figshare offers an important alternative to arXiv for empirical
biologists. Furthermore, by allowing all types of content, figshare arguably
provides an archive for early results (e.g.: figures, lab presentations).
figshare also supports commenting on preprints.

@@

F1000Research is not a public preprint server like the previous three servers.
Whereas arXiv, figshare, and PeerJ offer an option to submit a manuscript
without having it reviewed, papers submitted to F1000Research will eventually be
reviewed. Thus, F1000Research offers a hybrid model with publicly available
manuscripts at time of submission and standard peer-reviews. Manuscripts are
considered ``accepted'' and will only be indexed after two positive referee
response. F1000Research works closely with data providers to integrate raw
data to the paper. For instance, upon submitting a paper, authors are asked to
upload their data, which are then integrated in \emph{e.g.} figshare widgets,
the DOI of which are given in the paper when the data are first mentionned.
The licensing of the data is similar to that used by FigShare, meaning that
the articles are free to access, and can be redistributed readily. By
integrating data to the paper, F1000Research is working to make
science more reproducible and open.
